Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2023-06-30 manuscript.tex

\textcolor{blue}{Up until now, we specified models without a correlation between the individual effects. This means, for instance, that we do not expect that the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related. However, in cases where we would expect the estimation of individual effects to be related, we could estimate the correlation coefficient between individual effects. This would add complexity to the model. For the sake of the tutorial, we will focus on the models without individual effect correlation, but we illustrate the estimation of the model with correlation between the individual effects in} \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement C and D}.

@@

\includegraphics[width=1\linewidth]{I - Images/Fig4_distrtypefigure_v3} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig5_priorpredfigure_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig8_traceplots3pnew_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig9_posteriormu_v3} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig10_randompar_v2} \hfill{}

@@

\includegraphics[width=0.75\linewidth]{I - Images/Fig11_modelestimatefigure2_onlyrstan_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig12_traceplots3plog_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig13_posteriormulog_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig14_randomparlog_v2} \hfill{}

@@

\includegraphics[width=0.75\linewidth]{I - Images/Fig15_inddiflogplot_onlyrstan_v2} \hfill{}

@@

\includegraphics[width=0.6\linewidth]{I - Images/Fig16_sdrplotside_v2} \hfill{}

@@

\includegraphics[width=0.7\linewidth]{I - Images/Fig17_normalbfplot01} \hfill{}

@@

\textcolor{blue}{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity. We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model. This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences. For scenario 2, we remove individual variability from the side, digit, and full model. This resulted in the following posterior model probabilities shown in Table} \ref{tab:postmodproblog}.

@@

The examples illustrate the benefits of comparing hierarchical models of varying complexity. While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary. As shown in Tables\} \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.

>>>>

\textcolor{blue}{Up until now, we specified models without a correlation between the individual effects. This means, for instance, that we do not expect the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related. However, in cases where we would expect the estimation of individual effects to be related, we could estimate the correlation coefficient between individual effects. This would add complexity to the model. For the sake of the tutorial, we will focus on the models without individual effect correlation, but we illustrate the estimation of the model with correlation between the individual effects in} \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement C and D}.

@@

\includegraphics[width=1\linewidth]{I - Images/Fig5_distrtypefigure_v3} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig6_priorpredfigure_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig9_traceplots3pnew_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig10_posteriormu_v3} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig11_randompar_v2} \hfill{}

@@

\includegraphics[width=0.75\linewidth]{I - Images/Fig12_modelestimatefigure2_onlyrstan_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig13_traceplots3plog_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig14_posteriormulog_v2} \hfill{}

@@

\includegraphics[width=1\linewidth]{I - Images/Fig15_randomparlog_v2} \hfill{}

@@

\includegraphics[width=0.75\linewidth]{I - Images/Fig16_inddiflogplot_onlyrstan_v2} \hfill{}

@@

\includegraphics[width=0.6\linewidth]{I - Images/Fig17_sdrplotside_v2} \hfill{}

@@

\includegraphics[width=0.7\linewidth]{I - Images/Fig18_normalbfplot01} \hfill{}

@@

\textcolor{blue}{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity. We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model. This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation 8, specifying large individual differences. For scenario 2, we remove individual variability from the side, digit, and full model. This resulted in the following posterior model probabilities shown in Table} \ref{tab:postmodproblog}.

@@

The examples illustrate the benefits of comparing hierarchical models of varying complexity. While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary. As shown in Tables \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
